### 2019-06-08 paper/paper.tex
paper template

\documentclass[journal]{vgtc}                % final (journal style)
%\documentclass[review,journal]{vgtc}         % review (journal style)
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
%\documentclass[electronic,journal]{vgtc}     % electronic version, journal

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% This turns references into clickable hyperlinks.
\usepackage[bookmarks,backref=true,linkcolor=black]{hyperref} %,colorlinks
\hypersetup{
  pdfauthor = {},
  pdftitle = {},
  pdfsubject = {},
  pdfkeywords = {},
  colorlinks=true,
  linkcolor= black,
  citecolor= black,
  pageanchor=true,
  urlcolor = black,
  plainpages = false,
  linktocpage
}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Pleiades: Interactive Composing Tools for Vega-Lite Charts}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Chanwut Kittivorawong, Manesh Jhawar, Sorawee Porncharoenwase}
\authorfooter{
%% insert punctuation at end of each item
\item
 Chanwut Kittivorawong is an undergraduate Computer Science student at the University of Washington. E-mail: chanwutk@cs.washington.edu.
\item
 Manesh Jhawar is an undergraduate Computer Science student at the University of Washington. E-mail: mj06@uw.edu.
\item
 Sorawee Porncharoenwase is a graduate Computer Science student at the University of Washington. E-mail: sorawee@cs.washington.edu.
}

%other entries to be set up for journal
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
\abstract{
[...need better abstract here...]

Since Vega-Lite is a high-level grammar, the grammar is easy to understand,
so the range of users using the software can be from computer science expert to coding
beginner. Although users with less JSON experience can work with most of the Vega-Lite
easily, View Composition requires a good understanding of tree structure since View
Composition can be nested inside each other to create complex views. Beginner users
can struggle using this feature.

This software provides a graphical user interface for users to compose Vega-Lite charts.
The users can add charts that they want to work with to the software, then they can compose
them with layer, concat, repeat, and facet. They can also compose composed chart with one
of the four technique. The software also provide warnings when the users are trying to
compose charts that are incompatible with each other to avoid unexpected behaviors.
} % end of abstract

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Data Visualization, interactive system, Vega-Lite}

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

\CCScatlist{ % not used in journal version
 \CCScat{K.6.1}{Management of Computing and Information Systems}%
{Project and People Management}{Life Cycle};
 \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
}

%% Uncomment below to include a teaser figure.
%   \teaser{
%  \centering
%  \includegraphics[width=16cm]{screenshot.png}
%   \caption{add some picture?}
%   }

%% Uncomment below to disable the manuscript note
%\renewcommand{\manuscriptnotetxt}{}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

... intro here

% how to cite
facilisi~\cite{notes2002}

\section{Related Work}

Pleiades is built as a tool to help Vega-Lite users easily create multiple views
visualization by providing a Graphical User Interface to perform View Composition.

\subsection{Vega-Lite: A Grammar of Interactive Graphics}

...about Vega-Lite relating to this project...
% how to do foot note
\footnote{how to do foot note}.

\section{Methods}
\subsection{User Interface}

To work with Pleiades, users can add Vega-Lite specs that they are working with
to the left sidebar by clicking "NEW SPEC", then type in the Vega-Lite spec, and
save.

To create view composition, users can select view(s) as operand(s) and then apply
an operation. It will output the composed view in the main view area. Users are
allowed to select up to one view from the sidebar and up to one view from the main
view to perform an operation. Users can also perform an operation to composed views.
For example, a layered view can then be horizontal concatenated with another view.
Then, inside the concatenated views, the right view can be selected to repeat.

To edit composed view, users can select a view in the main view. Then, click "EDIT"
configure properties of the selected view.

Pleiades also perform minor validation before committing to any operation to prevent
an unexpected output. For example, when the user selects operands to layer, if one
of the operands is not a unit spec or layer spec, the "LAYER" button is disabled.
And if both operands do not have compatible axes, the "LAYER" button will show a
warning sign with tooltip that the axes are not compatible.

Pleiades also provides Inner View Navigator that shows the view from inner spec
of repeat or facet view. Repeat and facet operation produces a view containing
replication of the operand. When selecting repeat or facet view, the Inner View
Navigator shows the original view before the replication. This is useful when
the original view is also a composite view. Then, we can select the inner view
to edit from the Inner View Navigator.

Finally when the user is done with composing view, they can export to Vega-Lite
json file to use normally with any Vega-Lite compiler.

\subsection{Operations for View Composition}
There are 5 main operations users can do to compose views

\subsubsection{Place}
When the main view is empty, user can select a view in the sidebar. Then click
"PLACE" to place the view to the main view.

\subsubsection{Layer}

% how to do paragraph
\paragraph{how to add paragraph}

this is how to add pagrgraph

%% if specified like this the section will be committed in review mode
\acknowledgments{
The authors wish to thank A, B, C. This work was supported in part by
a grant from XYZ.}

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{template}
\end{document}

### 2019-06-08 src/variables.ts
hover color, mouse icon, fix bubbling

@@ -4,3 +4,8 @@ export const defaultVegaLiteHeight = 200;
 export const defaultVegaLiteWidth = 200;
 export const defaultVegaLiteShortHeight = 20;
 export const defaultVegaLiteShortWidth = 20;
+
+export const borderColor = 'lightgrey';
+export const hoverBorderColor = 'darkgrey';
+export const activeBorderColor = '#ed553b';
+export const mouseDownBorderColor = 'grey';

### 2019-06-09 src/SyntaxTree/LayerView.ts
allow layering with layer spec

@@ -28,11 +28,7 @@ export class LayerView extends CompositeView<UnitView> {
   public isCompatible(view: UnitView): boolean {
     const spec = view.export();
     return (
-      !spec['spec'] &&
-      !spec['layer'] &&
-      !spec['concat'] &&
-      !spec['hconcat'] &&
-      !spec['vconcat']
+      !spec['spec'] && !spec['concat'] && !spec['hconcat'] && !spec['vconcat']
     );
   }
 

### 2019-06-11 paper/paper.tex
add text to latex

@@
 \abstract{
-[...need better abstract here...]
+Vega-Lite is a high-level grammar which is easy to understand. Naturally,
+the user base of Vega-Lite ranges from beginners to experts in the field of
+visualization. Although users with less JSON experience have no trouble working
+with most of the features of Vega-lite, composing different views together has
+a sharp learning curve. View composition requires a good understanding of tree
+structure since different views can be nested inside each other to create more
+complex views.
 
-Since Vega-Lite is a high-level grammar, the grammar is easy to understand,
-so the range of users using the software can be from computer science expert to coding
-beginner. Although users with less JSON experience can work with most of the Vega-Lite
-easily, View Composition requires a good understanding of tree structure since View
-Composition can be nested inside each other to create complex views. Beginner users
-can struggle using this feature.
-
-This software provides a graphical user interface for users to compose Vega-Lite charts.
-The users can add charts that they want to work with to the software, then they can compose
-them with layer, concat, repeat, and facet. They can also compose composed chart with one
-of the four technique. The software also provide warnings when the users are trying to
-compose charts that are incompatible with each other to avoid unexpected behaviors.
+Pleiades provides a graphical user interface for users to compose Vega-Lite charts.
+The users can add charts that they want to work with the software, then use them to
+compose complex views. The operations they can perform are layer, concat, repeat, and
+facet. Using these four techniques, the users can keep on composing and constructing
+complex visualizations. Pleiades provides the users with warnings and restrictions when
+composing charts that would be incompatible. We provide this level of abstraction to
+help users focus more on the visualization and less about the nitty-gritty details of Vega-lite. 
 } % end of abstract
@@
 
-... intro here
+While working with Vega-Lite over the quarter, we realized that composing
+different charts had a learning curve. Users had to remember graph and data
+compatibility, code it in JSON, and if it was a complex composition, the JSON
+ended up being very nested and complicated. However, composing different graphs
+together is one of the heavily used features in the field of visualization, and
+having to deal with so much difficulty mentioned above seems like a hindrance to
+learning and visualizing data effortlessly. When we think about the users of
+Vega-lite, we have a predefined notion that the users are well learned in the
+field of data and computer science. However, with the recent growth and popularity
+in the data world, the user base for Vega-lite, now also comprises of entry-level
+data and visualization enthusiast and folks just interested in visualization without
+any coding background. Naturally, for beginners, working with heavily nested JSON
+specs can be dreadful.
 
-% how to cite
-facilisi~\cite{notes2002}
+Although Vega-Lite aims to be easy to use for users with a non-computer science
+background, view composition is one of the aspects that can be difficult to work
+with. Since Vega-Lite~\cite{notes2002} specification is in JSON format, when users
+want to nest view composition, inner specifications have to be nested heavily.
+However, mapping from the design in the user’s mind to JSON nested structure is
+hard because JSON does not well represent the layout of the composite views.
+
+Furthermore, when users design composite views, it makes sense for users to build
+composite views starting from a unit spec. Then, they can add more views and is
+able to experiment with the design by adding in and taking out views. With the
+nature of JSON, however, when working with Vega-Lite, specification is created
+from the outer most composite view and add inner views later. Additionally, when
+users need to make changes to the layout of composite views, they have to make
+changes to the JSON specification. With this, the users need to focus on how to
+implement the JSON, which would distract the focus of the user on the design.
+
+Pleiades is a toolkit for Vega-lite that gives the user the ability to compose
+charts without having to deal with remembering the rules of composition or working
+with the JSON. We provide a Graphical User Interface for users to add different
+pre-existing Vega-Lite specs and they use them to create complex compositions.
+We provide four options for composition: Layer, Concat, Repeat, and Facet. We also
+provide the users with an abstraction that handles all the rules for composition,
+by simply disabling the options when they cannot be performed. Pleiades has resulted
+in not only providing an efficient and easy way to create compositions but also
+enables the user to play with the data more efficiently.
 
@@
 
-Pleiades is built as a tool to help Vega-Lite users easily create multiple views
-visualization by providing a Graphical User Interface to perform View Composition.
+\subsection{Vega-Lite: A Grammar of Interactive Graphics}
+Vega-Lite is a high-level grammar for creating visualizations. The grammar abstracts
+away the low-level details of mapping data from dataset to the actual pixel of the
+working screen.
 
-\subsection{Vega-Lite: A Grammar of Interactive Graphics}
+Pleiades is built on top of Vega-Lite to further abstracts away the implementation
+details of view composition using JSON. In this work, we assume that users already
+are familiar with basic Vega-Lite that they are comfortable working with unit spec
+using Vega-Lite. Having all the unit specs the users need to compose, users can use
+Pleiades to compose them with the interactive GUI.
 
-...about Vega-Lite relating to this project...
 % how to do foot note
-\footnote{how to do foot note}.
+% \footnote{how to do foot note}.
 
@@
 
-To work with Pleiades, users can add Vega-Lite specs that they are working with
-to the left sidebar by clicking "NEW SPEC", then type in the Vega-Lite spec, and
-save.
+To work with Pleiades, users can firstly add Vega-Lite specs that they are working
+with to the left sidebar by clicking "NEW SPEC", then type in the Vega-Lite spec,
+and then save.
 
 To create view composition, users can select view(s) as operand(s) and then apply
-an operation. It will output the composed view in the main view area. Users are
-allowed to select up to one view from the sidebar and up to one view from the main
-view to perform an operation. Users can also perform an operation to composed views.
-For example, a layered view can then be horizontal concatenated with another view.
-Then, inside the concatenated views, the right view can be selected to repeat.
+an operation by clicking one of the operations in the operations bar on the top of
+the application. It will perform the operation from the selected view in the sidebar
+to the selected view in the working area. Users are allowed to select up to one view
+from the sidebar and up to one view from the working area to perform an operation.
 
-To edit composed view, users can select a view in the main view. Then, click "EDIT"
-configure properties of the selected view.
+To create more complex view composition, users can perform operations to the view
+in the working area. For example, a layered view can then be horizontally concatenated
+with another view. Then, inside the concatenated views, the right view can be selected
+to perform the repeat operation.
 
-Pleiades also perform minor validation before committing to any operation to prevent
-an unexpected output. For example, when the user selects operands to layer, if one
-of the operands is not a unit spec or layer spec, the "LAYER" button is disabled.
-And if both operands do not have compatible axes, the "LAYER" button will show a
-warning sign with tooltip that the axes are not compatible.
+To edit a composed view, users can select a view in the working area. Then, click
+"EDIT" configure the properties of the selected view.
 
-Pleiades also provides Inner View Navigator that shows the view from inner spec
-of repeat or facet view. Repeat and facet operation produces a view containing
-replication of the operand. When selecting repeat or facet view, the Inner View
-Navigator shows the original view before the replication. This is useful when
-the original view is also a composite view. Then, we can select the inner view
-to edit from the Inner View Navigator.
+Users can always redo or undo any action if they make mistakes by clicking "UNDO"
+or "REDO".
 
-Finally when the user is done with composing view, they can export to Vega-Lite
-json file to use normally with any Vega-Lite compiler.
+Pleiades also provides Inner View Navigator that shows the inner view of repeat
+or facet view. Repeat and facet operation produces a view containing replication
+of the inner view. When selecting a repeat or facet view, the Inner View Navigator
+shows the original view before the replication. This functionality is useful when
+the original view is also a composite view. Then, we can select the inner view to
+edit it from the Inner View Navigator.
+
+Finally, when the user is done with composing view, they can export the view in
+the working area to Vega-Lite JSON file to use normally with any Vega-Lite compiler.
 
@@
 \subsubsection{Place}
-When the main view is empty, user can select a view in the sidebar. Then click
-"PLACE" to place the view to the main view.
+When the working area is empty, user can select a view in the sidebar. Then click
+“PLACE” to place the view to the working area.
 
 \subsubsection{Layer}
+Users can perform layering when there are one selected view from the sidebar and
+one selected view from the working area. When clicking “LAYER”, the users are
+prompted with an option to layer the view in the sidebar over or under the view
+in the working area. Then, the selected view in the working area is layered with
+the selected view in the sidebar.
 
-% how to do paragraph
-\paragraph{how to add paragraph}
+\paragraph{Validation} Layering in Pleiades has some restrictions. First, both
+operands for layering has to be either unit views or layer views because
+Vega-Lite only allows layering unit and layer specs. Pleiades also perform an
+additional check axes compatibility of operands. Then, It will give a warning
+in the operation button to let the user know that layering can be done between
+these two operands but not fully compatible.
 
-this is how to add pagrgraph
+\paragraph{Editing} Users can edit a layer view by selecting the view in the
+working area. Then, click “EDIT”. The supported editing functionalities for layer
+view are to remove inner views and rearrange the layering order.
+
+\subsubsection{Concatenation}
+The user can perform concatenation when there is one selected view from the sidebar
+and one selected view from the working area. When clicking “CONCAT”, the user is
+prompted with an option to concatenate the selected view in the sidebar to the
+left, right, top, or bottom of the selected view in the working area.
+
+\paragraph{Editing} Users can edit a concat view by selecting the view in the
+working area. Then, click “EDIT”. In the edit popup window, the user can rearrange
+the order of concatenation, remove inner views, and switch between vertical and
+horizontal concatenation.
+
+\subsubsection{Facet}
+The user can perform faceting when there is one selected view from the working
+area. When clicking “FACET”, a popup window will show prompting the user to select,
+for each repeating direction (row or column or both), a field and its type to facet.
+Note that facetting will be applied to all inner views of the operand. Due to our time
+constraints, selecting a subset of inner views to only perform facet on the set of
+inner views is unavailable right now.
+
+\paragraph{Validation} Since parameters for facet are fields, facetting
+depends on the dataset. Thus, “FACET” button will show a warning signal when the
+operand has multiple data source. For instance, selecting a concat view that
+contains a chart from cars dataset and population dataset will result in having
+the warning signal in the “FACET”  button.
+
+\paragraph{Editing} There are two methods of modifying a facet view. Users can
+edit the properties of a facet view by selecting the editing view. Then, click
+“EDIT”. A popup window will show. Users can change the facet parameters for the
+facet view. An option for decompose facet view is also provided to replace the
+current facet view with its inner view (the original view). Another method for
+modifying a facet view is to modify the inner view of the facet view in the Inner
+View Navigator. When selecting a facet view, the inner view of the facet view will
+show up in the Inner View Navigator. Then, the user can perform modification to the
+view in Inner View Navigator. This modification, however, does not perform any
+validation on the datasets, due to the time constraints. If a new view is added
+and might break the well-formedness of the state. In practice, this is not a problem
+because Pleiades supports undo and redo, so users can always go back to the previous
+state when the output is invalid resulting from this modification.
+
+\subsubsection{Repeat}
+Every interaction in repeat operation works the same way as it would work in
+facet operation with the exception of the operation button to “REPEAT” and the
+popup window to configure parameter to perform repeat. The popup window will prompt
+the user to input, for each repeating direction (row or column or both), a list of
+fields to repeat and encoding channel to repeat. Repeating will also be applied to
+all inner views of the operand as same as faceting.
+
+\paragraph{Validation} Repeat operation performs the same validation as the facet
+operation does.
+
+\paragraph{Editing} Users can edit repeat view with the same methods as editing facet view.
+
+\subsection{Well-Formedness}
+As mentioned, Pleiades will restrict users from performing some operations that
+make views incompatible. More broadly, our design goal is that the working area
+should be well-formed throughout user interactions, where the working area is
+well-formed if it can be exported into a Vega-Lite spec. This might sound difficult
+to achieve, but the induction principle gives us a way to decompose the goal into
+subtasks: we only need to make sure that every Pleiades’ operation preserves
+well-formedness. As the initial view is trivially well-formed, by the induction
+principle, it follows that the working area will be well-formed at any time.
+
+The benefit of ensuring well-formedness at every step is that users will never
+need to correct a mistake. This contrasts with a tool that allows an ill-formed
+state because users then need to find mistakes, which could potentially be anywhere,
+and then fix them to make the state become well-formed. Without a good error reporting
+system that accurately guides users to fix problems, this process would be very time
+consuming and frustrating to users. A good error reporting system, however, is
+notoriously difficult to achieve. By ensuring well-formedness at every step, we
+can greatly simplify the system while providing good user experience.
+
+To implement well-formedness preserving operations, we perform speculation for
+each operation to validate if the operation should be permitted or not. There
+are two design decisions that we could make here: one is allowing users to attempt
+the operation, which will immediately fail, and the other is disallowing users
+from attempting the operation in the first place. The first approach has the advantage
+that the error messages could be very descriptive, allowing users to understand
+why the operation is not allowed. However, this comes with the cost of a more
+confusing user interface. We thus opt for the second approach which disallows
+users from attempting the operation in the first place. To mitigate the problem
+that users might not understand why an operation is not allowed, we additionally
+insert short text to explain why the operation is disabled for some potentially
+confusing operations. For example, when the user selects operands to layer, if
+one of the operands is not a unit spec or layer spec, the “LAYER” button is
+disabled. And if both operands do not have compatible axes, the “LAYER” button
+will show a warning sign with a tooltip that the axes are not compatible.
+
+In practice, the above approach is proven to be too rigid. One example of
+unintended consequences of the above approach is that users can’t edit any
+specification at all, because, during the editing, it’s very likely that
+the text won’t be able to parse correctly into the JSON format. We thus
+separate operations into macro-operations and micro-operations, where a
+macro-operation is a group of micro-operations. We then relax our goal to
+only apply to macro-operations. That is, every macro-operation preserves
+well-formedness. For micro-operations that does not preserve well-formedness,
+the operation won’t affect the working area directly. Rather, it is confined
+inside a dialog box so that a batch of micro-operations can be readily reversed.
+
+\subsection{Syntax Tree}
+The state of the composite view for the output in the working area is internally
+stored as a tree. Each node of the tree represents a view composition, which has
+different properties depending on the type of composition.
+
+\subsubsection{UnitView} contains a Vega-Lite specification. However, Pleiades does
+not perform any validation to check if the specification is actually a Vega-Lite’s
+UnitSpec. So, UnitView can be layer, concat, repeat, or facet spec but cannot be
+modified by Pleiades’ operations.
+
+\subsubsection{LayerView} has a list of children nodes representing inner views that
+are being layered. The children are only allowed to be UnitView or LayerView.
+More specifically, the specification of the UnitView child has also to be either
+unit or layer spec as Vega-Lite only allows layering layer and unit specs.
+
+\subsubsection{ConcatView} has a list of children node representing inner views that
+are being concatenated and a type of concatenation to be either horizontal or
+vertical concatenation. The children of ConcatView can be any type of View.
+
+\subsubsection{FacetView} has one inner view and facet properties, which contains
+field and type of row facet and/or field and type of column facet. The inner
+view can be any type of Views. All the datasets in the inner view should come
+from the same source if the inner view is a composite view. Although Pleiades
+does not prevent users from facet-ing View that contains multiple data sources,
+it gives a warning to the users that the datasets are not compatible.
+
+\subsubsection{RepeatView} has one inner view and repeat properties, which contains
+a list of fields and encoding channel of row repeat and/or a list of fields and
+encoding channel of column repeat. Everything else of RepeatView is the same as
+FacetView.
+
+\paragraph{The benefit of using syntax tree} With the syntax tree, inner views
+can be located to modify easily. Furthermore, when the working view is output to
+the working area, it can be rendered recursively through each node in the tree.
+This recursive rendering allows us to render inner views as separate views from
+the outer view. For example, when rendering concat view, the inner views are
+rendered separately in a way that users can select either the inner view or
+the whole concat view.
+
+\subsection{How data of each spec is pulled to the top level in facet and repeat}
+In facet and repeat spec, data must be at the topmost level of the output JSON
+spec. When RepeatView or FacetView is exported to JSON, Pleiades pull out all
+the data source in every children view of the inner view to the top level. If
+there is more than one distinct data source, Pleiades will use the first one it
+encounters by pre-order tree traversal. Pleiades also uses a similar method to
+exhaustively search for every data source in the inner views of an operand to
+validate if there should be a warning shown for the current operand for facet/repeat.
+
+\section{Results}
+
+result here
+
+\section{Discussion}
+
+discussion here
+
+\section{Future Work}
+We are currently working on extending our software to perform a better validation
+of every operation. This validation is important for the software because the
+software can then maintain the well-formedness of the output at every state after
+every operation. In the current build of the software, we decided to left out some
+validations due to the time constraints. However, in the real scenario, not validating
+everything is not a problem since our software support undo and redo that users
+can always reverse back to the previous state.
+
+As observed from users using Pleiades, users try dragging view to the main view
+first when first trying the software. As an extension to Pleiades, we are looking
+forward to supporting dragging and dropping in the software to provide a more natural
+interaction to the users.
+
+When adding a new spec to the sidebar in the software, Pleiades interpret the new
+View as a UnitView regardless of its actual composite type. While this does not
+create a problem in practice, the rendered view is always static. For instance,
+the inner views of concat spec that is added by “NEW SPEC” are not selectable by
+users, and cannot be edited using the “EDIT” button.  Our future work for addressing
+this issue is that we would incorporate a parser that parses new specs to syntax tree
+of composite views so that all the functionality is supported for the composite
+views added by users.
 

### 2019-06-11 paper/paper.tex
use `` as starting double quote

@@
 To work with Pleiades, users can firstly add Vega-Lite specs that they are working
-with to the left sidebar by clicking "NEW SPEC", then type in the Vega-Lite spec,
+with to the left sidebar by clicking ``NEW SPEC'', then type in the Vega-Lite spec,
 and then save.
@@
 To edit a composed view, users can select a view in the working area. Then, click
-"EDIT" configure the properties of the selected view.
+``EDIT'' configure the properties of the selected view.
 
-Users can always redo or undo any action if they make mistakes by clicking "UNDO"
-or "REDO".
+Users can always redo or undo any action if they make mistakes by clicking ``UNDO''
+or ``REDO''.
 

### 2019-06-12 src/variables.ts
thicken border; add dropdown arrow; lighten gradient

@@ -1,5 +1,5 @@
 export const sidebarWidth = 150;
-export const borderWidth = 2;
+export const borderWidth = 4;
 export const defaultVegaLiteHeight = 200;
 export const defaultVegaLiteWidth = 200;
 export const defaultVegaLiteShortHeight = 20;

### 2019-06-12 src/global.d.ts
implement replace

@@ -32,7 +32,7 @@ interface IState {
   tree: View | null;
 }
 
-type Operator = 'layer' | 'concat' | 'repeat' | 'facet' | 'place';
+type Operator = 'layer' | 'concat' | 'repeat' | 'facet' | 'place' | 'replace';
 
 type InsertOrder = 'append' | 'prepend';
 

### 2019-06-12 src/store.ts

@@ -185,6 +185,14 @@ const reducer = (globalState = initialState, action: Action): IGlobalState => {
                   newTree = facet;
                 }
                 break;
+              case 'replace':
+                const leftView = new UnitView(findInNav(left[0]));
+                if (parent) {
+                  parent.replaceChild(leftView, view.id);
+                } else {
+                  newTree = leftView;
+                }
+                break;
               default:
                 return assertNever(action.operator);
             }

### 2019-06-12 src/variables.ts

@@ -1,5 +1,5 @@
 export const sidebarWidth = 150;
-export const borderWidth = 4;
+export const borderWidth = 3;
 export const defaultVegaLiteHeight = 200;
 export const defaultVegaLiteWidth = 200;
 export const defaultVegaLiteShortHeight = 20;

### 2019-06-12 final/paper/paper.tex
move poster and paper to final folder

--- 2019-06-11 paper/paper.tex
@@
 
-\subsubsection{Place}
+\subsubsection{Place/Replace}
 When the working area is empty, user can select a view in the sidebar. Then click
-“PLACE” to place the view to the working area.
+“PLACE” to place the view to the working area. When the working area is not empty,
+the same button's text changes to ``REPLACE''. Users can click this button when
+there are one selected view from the sidebar and one selected view from the working
+area. Then, the selected view in the working area is replaced with the selceted
+view in the sidebar.
 

### 2019-06-12 final/paper/paper.tex

@@
 nature of JSON, however, when working with Vega-Lite, specification is created
-from the outer most composite view and add inner views later. Additionally, when
+from the outermost composite view and add inner views later. Additionally, when
 users need to make changes to the layout of composite views, they have to make
@@
 
-Pleiades is built on top of Vega-Lite to further abstracts away the implementation
-details of view composition using JSON. In this work, we assume that users already
-are familiar with basic Vega-Lite that they are comfortable working with unit spec
-using Vega-Lite. Having all the unit specs the users need to compose, users can use
-Pleiades to compose them with the interactive GUI.
+Pleiades is built on top of Vega-Lite to further abstract away the implementation
+details of view composition using JSON. In this work, we assume that users are
+already familiar with basic Vega-Lite that they are comfortable working with unit
+spec using Vega-Lite. Having all the unit specs the users need to compose, users
+can use Pleiades to compose them with the interactive GUI.
 

### 2019-06-12 final/paper/paper.tex
Add results

--- 2019-06-12 final/paper/paper.tex
@@
 
+One more results that we found was, having the ability to perform the four operations: 
+Layer, Concat, Facet and Repeat, it gave user a better understanding of these 
+operations and how to properly use then. The validation steps, along with the 
+error message that Pleiades provides helps user to understand what they are doing 
+wrong, helping them in their education about visualization in general.
+
+In terms of exploring data, the time take to compose different charts and see 
+find useful results, went down by a lot because of the high-level of interactivity 
+provided. Users don’t have to worry about updating JSON code for ever 
+small change in their chart.
+
+Overall, Pleiades resulted in easing beginners to develop composite charts,
+save time for users, ease data-exploration, and act as a support toolkit for education.
+
 \section{Results}
 
-result here
+The key result for Pleiades was how easy it became to compose graph. During our presentation, 
+we asked a lot of our peers to try out the our application and compare their workflow to that 
+when they used the Vega-Lite API along with a JSON spec. The results were as we expected: 
+Pleiades gives a better understanding and ease at composing different charts and see what fits the best.
+The key-idea behind why Pleiades makes composing charts easier and more intuitive is because the 
+process is bottom-up, whereas Vega-lite is top-down. Also, the graphical user interface provided 
+used with a smother way to compose charts, than to write nested JSON specs.
 

### 2019-06-12 final/paper/paper.tex
Grammar check results

--- 2019-06-12 final/paper/paper.tex
@@
 
-result here
+The key result for Pleiades was how easy it became to compose graph. 
+During our presentation, we asked a lot of our peers to try out our 
+application and compare their workflow to that when they used the 
+Vega-Lite API along with a JSON spec. The results were as we expected: Pleiades 
+gives a better understanding and ease at composing different charts and see
+what fits the best. The key-idea behind why Pleiades makes composing charts 
+easier and more intuitive is because the process is bottom-up, 
+whereas Vega-lite is top-down. Also, the graphical user interface 
+provided used with a smother way to compose charts, than to write nested JSON specs.
+
+One more results that we found was, having the ability to perform 
+the four operations: Layer, Concat, Facet and Repeat, it gave the user
+a better understanding of these operations and how to properly use then. 
+he validation steps, along with the error message that Pleiades provides 
+helps users to understand what they are doing wrong, helping them in their 
+education about visualization in general.
+
+In terms of exploring data, the time it takes to compose different 
+charts and see find useful results, went down by a lot because of the 
+high-level of interactivity provided. Users don’t have to worry about 
+updating JSON code for every small change in their chart.
+
+Overall, Pleiades resulted in easing beginners to develop composite 
+charts, save time for users, ease data-exploration, and act as a 
+support toolkit for education.
 

### 2019-06-12 final/paper/paper.tex
add screenshot

@@
 
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=3in]{replace.png}
+ \caption{Before and after replacing the selcted bar chart with the selected scatter plot}
+\end{figure}
+
 \subsubsection{Layer}
@@
 
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=3in]{layer.png}
+ \caption{Before and after layering the selcted error-band chart with the selected scatter plot}
+\end{figure}
+
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=3in]{layer_warn.png}
+ \caption{Warning when axes of layering views are not fully compatible (between the bar chart and the scatter plot)}
+\end{figure}
+
 \subsubsection{Concatenation}
@@
 
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=3in]{concat.png}
+ \caption{Before and after concatenating the selcted bar chart with the selected scatter plot}
+\end{figure}
+
 \subsubsection{Facet}
@@
 for each repeating direction (row or column or both), a field and its type to facet.
-Note that facetting will be applied to all inner views of the operand. Due to our time
+Note that facet-ing will be applied to all inner views of the operand. Due to our time
 constraints, selecting a subset of inner views to only perform facet on the set of
@@
 
-\paragraph{Validation} Since parameters for facet are fields, facetting
+\paragraph{Validation} Since parameters for facet are fields, facet-ing
 depends on the dataset. Thus, “FACET” button will show a warning signal when the
@@
 
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=3in]{facet.png}
+ \caption{Before and after facet-ing the selected scatter plot with ``Origin'' as the field for facet-ing column.}
+\end{figure}
+
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=3in]{facet_warn.png}
+ \caption{Warning on repeat and facet when not every dataset in the selected view are from the same source.}
+\end{figure}
+
 \subsubsection{Repeat}
@@
 \paragraph{Editing} Users can edit repeat view with the same methods as editing facet view.
+
+\begin{figure}[htb]
+ \centering
+ \includegraphics[width=3in]{repeat.png}
+ \caption{Before and after repeating the selected scatter plot with ``Displacement'' and ``Cylinders'' as the list of fields for repeat column on x-axis.}
+\end{figure}
 
@@
 
-The key result for Pleiades was how easy it became to compose graph. 
-During our presentation, we asked a lot of our peers to try out our 
-application and compare their workflow to that when they used the 
-Vega-Lite API along with a JSON spec. The results were as we expected: Pleiades 
-gives a better understanding and ease at composing different charts and see
-what fits the best. The key-idea behind why Pleiades makes composing charts 
-easier and more intuitive is because the process is bottom-up, 
-whereas Vega-lite is top-down. Also, the graphical user interface 
-provided used with a smother way to compose charts, than to write nested JSON specs.
-
-One more results that we found was, having the ability to perform 
-the four operations: Layer, Concat, Facet and Repeat, it gave the user
-a better understanding of these operations and how to properly use then. 
-he validation steps, along with the error message that Pleiades provides 
-helps users to understand what they are doing wrong, helping them in their 
-education about visualization in general.
-
-In terms of exploring data, the time it takes to compose different 
-charts and see find useful results, went down by a lot because of the 
-high-level of interactivity provided. Users don’t have to worry about 
-updating JSON code for every small change in their chart.
-
-Overall, Pleiades resulted in easing beginners to develop composite 
-charts, save time for users, ease data-exploration, and act as a 
-support toolkit for education.
+result here
 

### 2019-06-12 final/paper/paper.tex
+Result +Discussion

--- 2019-06-12 final/paper/paper.tex
@@
 the operation won’t affect the working area directly. Rather, it is confined
-inside a dialog box so that a batch of micro-operations can be readily reversed.
+inside a dialog box so that a batch of micro-operations can be readily reverted.
 
@@
 
-\paragraph{The benefit of using syntax tree} With the syntax tree, inner views
-can be located to modify easily. Furthermore, when the working view is output to
+\paragraph{The benefit of syntax tree} With the concept of syntax tree, inner views
+can be located to modify easily. When the working view is output to
 the working area, it can be rendered recursively through each node in the tree.
@@
 
-\subsection{How data of each spec is pulled to the top level in facet and repeat}
+\subsection{Data Pulling}
 In facet and repeat spec, data must be at the topmost level of the output JSON
@@
 
-result here
+To evaluate our system, we perform a case study aiming to answer the following
+questions:
+
+\begin{itemize}
+\item How easy is it to use Pleiades?
+\item Does Pleiades allow effective view composition?
+\end{itemize}
+
+For the question how easy it is to use Pleiades, two participants of the case
+study report back that the tool is mostly easy to use. They wish that
+the tool provides more visual cues and that they prefer drag and drop to
+operand/operator selection, however. Some of these suggestions are already
+incorporated into the final version of the tool.
+
+For the question whether Pleiades allows effective view composition,
+the participants' feedback is highly positive, stating that charts
+can be composited in an intuitive way and visually responsive.
+
+During our poster presentation, we additionally asked a lot of our peers
+to try out our application and compare their workflow to that when they used the
+Vega-Lite API along with a JSON spec. The feedback is similar to what we have
+seen in the case study.
+
+Based on these results, we conclude that Pleiades eases beginners to
+develop composite charts, save time for users, ease data-exploration,
+and act as a support toolkit for education.
 
@@
 
-discussion here
+As we expected, the results show that Pleiades gives users a better understanding
+and ease them at composing different charts and see what fits the best.
+The key idea behind why Pleiades makes composing charts
+easier and more intuitive is because the process allows users to compose small
+charts and rearrange them into a bigger charts. We call this approach
+\emph{bottom-up}. This contrasts with Vega-lite's \emph{top-down approach}
+which demands users to know the layout of the entire structure already
+beforehand and hence is very rigid and inflexible. Also, the graphical user interface
+provided used with a smoother way to compose charts, than to write nested JSON specs.
+
+One interesting result we found was that the ability to perform
+the four operations gave users a better understanding of these operations
+and how to properly use then. The validation steps, along with the error message
+that Pleiades provides helps users to understand what they are doing wrong,
+helping them in their education about visualization in general.
+
+In terms of exploring data, the time it takes to compose different
+charts and see find useful results, went down by a lot because of the
+high-level of interactivity provided. Users don’t have to worry about
+updating JSON code for every small change in their chart. Moreover,
+a bad layout could be a hindrance that shows interesting trends in a chart.
+Our tool which allows users to experiment with composition easily thus helps
+with data exploration too.
 

### 2019-06-12 final/paper/paper.tex
+bib system

@@
 \abstract{
-Vega-Lite is a high-level grammar which is easy to understand. Naturally,
+  Vega-Lite~\cite{2017-vega-lite} is a high-level grammar which is easy to understand. Naturally,
 the user base of Vega-Lite ranges from beginners to experts in the field of
@@
 background, view composition is one of the aspects that can be difficult to work
-with. Since Vega-Lite~\cite{notes2002} specification is in JSON format, when users
+with. Since Vega-Lite specification is in JSON format, when users
 want to nest view composition, inner specifications have to be nested heavily.
@@
 %\nocite{*}
-\bibliography{template}
+\bibliography{paper.bib}
 \end{document}

### 2019-06-12 final/paper/paper.tex
typo

@@
 \item
- Manesh Jhawar is an undergraduate Computer Science student at the University of Washington. E-mail: mj06@uw.edu.
+ Manesh Jhawar is an undergraduate Computer Science student at the University of Washington. E-mail: mj06@cs.washington.edu.
 \item
@@
 easier and more intuitive is because the process allows users to compose small
-charts and rearrange them into a bigger charts. We call this approach
+charts and rearrange them into a bigger chart. We call this approach
 \emph{bottom-up}. This contrasts with Vega-lite's \emph{top-down approach}
@@
 beforehand and hence is very rigid and inflexible. Also, the graphical user interface
-provided used with a smoother way to compose charts, than to write nested JSON specs.
+provided users with a smoother way to compose charts, than to write nested JSON specs.
 
@@
 the four operations gave users a better understanding of these operations
-and how to properly use then. The validation steps, along with the error message
+and how to properly use them. The validation steps, along with the error message
 that Pleiades provides helps users to understand what they are doing wrong,
@@
 %% if specified like this the section will be committed in review mode
-\acknowledgments{
-The authors wish to thank A, B, C. This work was supported in part by
-a grant from XYZ.}
+% \acknowledgments{
+% The authors wish to thank A, B, C. This work was supported in part by
+% a grant from XYZ.}
 
